Sketching for Density Estimation

### 2018-12-21 komlos.tex
gamma_2

\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}{Conjecture}[section]
%\newtheorem{claim}[theorem]{Claim}
\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{remark}[theorem]{Remark}
%\newtheorem{definition}[theorem]{Definition}
%\newtheorem{observation}[theorem]{Observation}
%\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{fact}[theorem]{Fact}

@@

\newcommand{\ip}[1]{\left \langle #1 \right \rangle}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\D}{\rho}
\newcommand{\eps}{\epsilon}
\newcommand{\F}{\mathcal{F}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Q}{\mathcal{Q}}
%\usepackage{ntheorem}
\newcommand{\floor}[1]{\left \lfloor #1 \right \rfloor}
\newcommand{\ceil}[1]{\left \lceil #1 \right \rceil}

@@

@@

%
%\subsection{notes about row selection}
%The problem is 
%$$ \min_s \| \sum_i \sigma_i x_ix_i^T\| $$
%where $\|x_i\| \leq 1$. Consider the following greedy approach: At point $x_i$, let
%$$ C_i = \sum_{j<i} \sigma_j x_j x_j^T, \ \  l_+ = \|C_i + x_i x_i^T \|, \ l_- = \|C_i - x_i x_i^T\|$$
%We choose $\sigma_i = 1$ if $l_+ < l_i$, otherwise $\sigma_i=-1$.
%
%This approach can be shown to be better than random. Random will result in an answer of up to 
%
%Consider the following example. 
%$$x_1=e_1, x_2=e_2*\eps, x_3=e_2, x_4=e_1/\sqrt{2} + e_2/\sqrt{2}$$
%where $\eps \to 0$. We get that the covariance matrix after inserting $x_3$ has two eigenvalues of +1 and -1. Now when inserting $x_4$ we have to increase the magnitude of one of them and we exceed the limit of 1. 
%
\bibliographystyle{plain}
\bibliography{density}

>>>>

\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}{Conjecture}[section]
%\newtheorem{claim}[theorem]{Claim}
\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{remark}[theorem]{Remark}
%\newtheorem{definition}[theorem]{Definition}
%\newtheorem{observation}[theorem]{Observation}
\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{fact}[theorem]{Fact}

@@

\newcommand{\ip}[1]{\left \langle #1 \right \rangle}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\D}{\rho}
\newcommand{\eps}{\epsilon}
\newcommand{\F}{\mathcal{F}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Q}{\mathcal{Q}}
%\usepackage{ntheorem}
\newcommand{\floor}[1]{\left \lfloor #1 \right \rfloor}
\newcommand{\ceil}[1]{\left \lceil #1 \right \rceil}
\newcommand{\disc}{\text{disc}}

@@

A useful fact about the $\gamma_2$ norm is that it is indeed a norm. This is proven in \cite{tomczak1989banach}. In \cite{matouvsek2014factorization} this is listed in Section 2, Proposition 2.2 along with other known properties of the $\gamma_2$ norm.

@@

%
%\subsection{notes about row selection}
%The problem is 
%$$ \min_s \| \sum_i \sigma_i x_ix_i^T\| $$
%where $\|x_i\| \leq 1$. Consider the following greedy approach: At point $x_i$, let
%$$ C_i = \sum_{j<i} \sigma_j x_j x_j^T, \ \  l_+ = \|C_i + x_i x_i^T \|, \ l_- = \|C_i - x_i x_i^T\|$$
%We choose $\sigma_i = 1$ if $l_+ < l_i$, otherwise $\sigma_i=-1$.
%
%This approach can be shown to be better than random. Random will result in an answer of up to 
%
%Consider the following example. 
%$$x_1=e_1, x_2=e_2*\eps, x_3=e_2, x_4=e_1/\sqrt{2} + e_2/\sqrt{2}$$
%where $\eps \to 0$. We get that the covariance matrix after inserting $x_3$ has two eigenvalues of +1 and -1. Now when inserting $x_4$ we have to increase the magnitude of one of them and we exceed the limit of 1. 
%

\section{Active Learning for linear functions (based on $\gamma_2$ norm)}

Let $x_1,\ldots,x_n \in R^d$ be and $y_1,\ldots,y_n$ be labels associated to the data points. Let $K$ be a psd kernel. Consider a problem of learning a linear function aiming to minimize the loss of 
$$\min_w \sum_{i=1}^n L(K(x_i, w), y_i)$$

For a function $f(q,x)$, let $A_f$ be a matrix whose rows are an $\eps$-net over the Euclidean unit ball of $R^d$, and columns correspond to the different $x_i$'s. The value of $A_f$ at an entry corresponding to $(q,x)$ if $f(q,x)$.

\begin{definition}
For a matrix $A$,
$$ \|A\|_{\gamma_2} = \min_{B,C: A=BC^T} \|B\|_{2 \to \infty} \|C\|_{2 \to \infty}$$
is the $\gamma_2$ norm of $A$. 
\end{definition}
 \cite{tomczak1989banach} prove that this is indeed a norm, in particular that 
 $\|A_1+A_2\|_{\gamma_2} \leq \|A_1\|_{\gamma_2} + \|A_2\|_{\gamma_2}$.

\begin{lemma} \label{lem:powers}
For $k \geq 0$ let $f_k(q,x) = \ip{q,x}^k$. If holds that $\|A_{f_k}\|_{\gamma_2} \leq 1$
\end{lemma}
\begin{proof}
For $k=0$ the statement is obvious. For $k>0$, notice that $f_k$ is a kernel, and in particular there exists a mapping $\phi_k: \R^d \to \R^{d^k}$ such that
$$ \ip{q,x}^k = \ip{\phi_k(q), \phi_k(x)} $$
Furthermore, 
$$ \|\phi_k(x)\| = \sqrt{\ip{\phi_k(x), \phi_k(x)}} = \sqrt{\|x\|^{2k}} = \|x\|^k \leq 1 $$
Since all $x,q$ have at most unit norm we get a bound over the $\gamma_2$ norm by viewing the $B,C$ matrices obtained from the $\phi_k$ embeddings of the different $x,q$'s
\end{proof}

\begin{lemma}
Let $L:\R \to \R$ be a function that can be written as 
$$ L(z) = \sum_{k=0}^\infty \alpha_k z^k$$
with $\sum_{k=0}^{\infty} |\alpha_k| \leq C$. For $f_L(q,x) = L(\ip{q,x})$ it holds that 
$$ \|A_{f_L}\|_{\gamma_2} \leq C $$
\end{lemma}
\begin{proof}
This is a direct corollary from $\gamma_2$ being a norm, and Lemma~\ref{lem:powers}
\end{proof}

This lemma in particular holds for the logistic function as its Taylor series has elements that converge. In particular, evaluating the function at $i$ and summing the real and imaginary parts of the result gives the sum of the absolute values of the coefficients, which is (roughly) $0.461$
\begin{corollary}
For $L(z) = \log(1+e^z)$ we have $\|A_{f_L}\|_{\gamma_2} \leq 0.5$
\end{corollary}

\begin{definition}
For a matrix $A$ of dimensions $m \times n$, $\disc(A)$ is defined as 
$$ \min_{s \in \{-1,1\}^n} \|As\|_\infty $$
\end{definition}
In \cite{matouvsek2014factorization} they prove that 
$$\|A\|_{\gamma_2} / \log(m) \leq \disc(A) \leq \|A\|_{\gamma_2}  \sqrt{\log(m)}$$
This gives a way to control the discrepancy of a matrix by representing it as a sum of matrices with known discrepency or $\gamma_2$ norm. In particular, for the logistic function we get that 
$$\disc(A_{f_L}) \leq \sqrt{\log(|Q|)} $$
where $Q$ is an $\eps$-net over the Euclidean unit sphere. This gives the bound of 
$$\disc(A_{f_L}) \leq O(\sqrt{d\log(1/\eps)}/\eps) $$

Let's discuss what this means. For a logistic regression problem, by observing the weights and labels we can change the signs of the inputs so that the logistic loss is always $\log(1+\exp(\ip{q,x}))$, and we can now convert the discrepancy result into a coreset of size $\sqrt{d\log(1/\eps)}/\eps$, as opposed to a random sample of $1/\eps^2$.

Let $K$ be a psd kernel, meaning for $x,q \in \R^d$ we have for some mapping $\phi$ into a Hilbert space $K(q,x) = \ip{\phi_K(q), \phi_K(x)}$. Notice that for $\phi_k$ defined above
$$ \ip{\phi_k(\phi_K(q)), \phi_k(\phi_K(x))} = \ip{\phi_K(q), \phi_K(x)}^k = K(q,x)^k $$
meaning that $K^k(q,x)$ is a psd kernel for all $k \geq 0$. By using the exact same arguments before we can define $A_{K,k} \in \R^{|Q|,n}$ with $A_{K,k}(q,i) = K(q, x_i)^k$ and conclude that $\|A_{K,k}\|_{\gamma_2} \leq 1$. Then, for the matrix $A_{K,f_L}$ with
$$ L(z) = \sum_{k=0}^\infty \alpha_k z^k, A_{K,f_L}(q,i) = L(K(q,x_i))$$
we have 
$$\|A_{K,f_L}\|_{\gamma_2} \leq \sum_{k=0}^{\infty} |\alpha_k| $$

\begin{corollary}
For $L(z)=\log(1+\exp(z))$ and $L(z) = z^2$, and any psd kernel $K:\R^d \times \R^d \to \R$ we have
$$\|A_{K,f_L}\|_{\gamma_2} \leq 1 $$ 
and
$$\disc(A_{K,f_L}) \leq O(\sqrt{d\log(1/\eps)}/\eps) $$ 
\end{corollary}

%Let's discuss what this means. For a logistic regression problem, by observing the weights we can change the signs of the inputs so that the logistic loss is always $\log(1+\exp(\ip{q,x}))$, and we can now convert the discrepancy result into a coreset of size $\sqrt{d\log(1/\eps)}/\eps$, as opposed to a random sample of $1/\eps^2$. This has a limitation where we need the labels to determine the coreset. We now proceed to see how we can remove that need.
%
%We will require an assumption regarding the labels of the data: We assume that for some $q \in Q$, there exist at most $\rho n$ data point label pairs $(x,y) \in X \times \{-1,1\}$ such that $\text{sign}(\ip{q,x}) \neq y$.

\bibliographystyle{plain}
\bibliography{density}
